Making the Best of a Bad Situation

### 2022-03-21 paper.tex
space hacked

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
    TODO
\end{abstract}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Motivation}
Property-Based Testing (PBT) \cite{qc} is a program testing method which tests
logical properties of functions by randomly generating thousands of inputs, and ensuring that
the the properties hold at each concrete value. While originally concieved as a more
principled form of unit testing, PBT also works as a lightweight formal verification method
wherein the specifications for a program are \emph{tested} instead of proved.

The main idea behind this approach to verification is that PBT trades off the
full certainty provided by SMT and theorem proving approaches for a significantly streamlined
user experience requiring considerably less user input. To ``verify'' a universally quantified property
$\forall x. P(x) \implies Q(x)$, we generate values satisfying $P$, and check that they satisfy $Q$.
Unfortunately PBT does still requre \emph{some} user input. Users must occasionally write \emph{generators}: programs which
emit random data that satisfies the premise $P$.

Much of the time, this burden can be mitigated by automation, such as when the data structure being generated
is composed entirely of structures that already have generators, or $P$ is described as an inductive
relation \cite{gggir}. However, the case of inferring generators for \emph{numerical} properties is not well
handled by prior work. Numerical properties---those involving constriants over numbers---appear regularly in
systems verification tasks \cite{???}, where structured data described by inductive invariants is rare, inputs regularly
take the form of pointers and integers, and many function preconditions amount to bounds-checking.

In this paper, we develop a method for inferring random generators for a small class of numerical properties akin
to the kind commonly found in systems verification. Because manual generator-writing for numerical properties is especially
tedious, our goal is to infer generators that are comparable to ones that a user might write as a first cut.
The method proceeds by developing a number of generator ``candidates'', expressed in a DSL we call \texttt{ALuck}.
These candidates mimic the way that a human might write a generator for numerical preconditions: by randomly choosing values
one variable at a time, constraining subsequent choices by the previous values chosen. As we will see, it is sometimes the
case that \emph{none} of the candidates are all that good. In this case, we must make the best of a bad situation, and
discover the best candidate among the set.

In \autoref{sec:luck}, we introduce the syntax of the properties we will handle,
as well as the syntax and semantics of \texttt{ALuck}. Next, in \autoref{sec:sci},
we present our algorithm for inferring candidate
generators for a precondition. Then in \autoref{sec:bandits}, we discuss the online
learning technique we employ to discover the best generator among our
candidates. Finally, in \autoref{sec:impl}, we describe our implementation of
this algorithm, present benchmarks, and briefly discuss future work.

@@

The syntax of propositions that we handle are shown in
\autoref{fig:precond-syntax}. These are conjunctions and negations of
arithmetical equalities and inequalities with variables, which range over
the arguments of the function under test. The allowable expressions in these inequalities are
essentially multi-linear functions, where each variable occurs with degree at
most one. While this form is restrictive, we have found empirically that this covers
a wide range of preconditions in systems verification. Moreover, there
appears to be no inherent difficulty to extending the technique to handle
other numerical operators (div, mod), and constraints over structured types like lists: we present the algorithm as-is to
simplify the discussion.

\begin{figure}
\caption{Syntax of Propositions}
\label{fig:precond-syntax}
$$
\begin{array}{llll}
\text{Expressions} & e & ::= & x \; | \; e + e \; | \; -e \;|\; e*e\\
\text{Relations} & R & ::= & = \;|\; \leq \;|\; <\\
\text{Propositions} & P & ::= & P \wedge P \;|\; e R e\\
\end{array}
$$
\end{figure}

In order to make the generator inference/synthesis problem tractable, we fix the
syntactic form of the generators we plan to consider. Our generators are written
in a language called \texttt{ALuck} (for \textit{arithmetic} \texttt{Luck})
inspired by \texttt{Luck} \cite{luck}, a domain-specific language for writing
generators which tries to mirror the declarative form of standard propositions.
Generators written \texttt{ALuck} run by sequentially \textit{constraining} and
\textit{concretizing} variables. Every variable in an \texttt{ALuck} generator
begins as a symbolic variable. Constraints over these variables are then added.
Variables can then be concretized, wherein they are replaced by a random value
satisfying the constraints on that variable that have been accumulated thusfar.
The final result of the generator is a map from variables to their (randomly)
chosen values.

More concretely, generators in \texttt{ALuck} are sequences of ``concretize"
operations, written $!x$, and ``constrain" operations, written simply as the
constraint to be added. The syntax of \texttt{ALuck} is shown in
\autoref{fig:aluck-syntax}. These sequences are then evaluated from left to
right while maintaining a pair of mappings: one from concretized variables to
their values, and the other from yet-unconcretized variables to the set of
constraints that have accumulated on them.  When a ``constrain" operation $c$ is
encountered, the constraint $c$ is added to the constraint sets of all of the
variables it mentions. If a constraint mentions no variables, it is checked for
valdity: if the constraint is not valid, the generator \emph{fails}.

When a ``concretize" operation $!x$ is encountered, a value is randomly sampled from
the uniform distribution on the set of possible values\footnote{ Because
integers are bounded by machine word lengths, the ``uniform distribution" does
make sense here, even for unconstrained variables.  } denoted by the constraints
on $x$. This semantics is shown in \autoref{fig:aluck-semantics}, where the
judgment $V,C \vdash s \Downarrow V',C'$ means that the script $s$ evaluates
under concretized-variable-map $V$ and constraint map $C$, and returns updated
variable and constraint maps $V'$ and $C'$.

Of course, the step of ``sampling from constraints specified by inequalities on
variables" for concretization operations is exactly what we're trying to solve!
To make this step tractable, we enforce that our \texttt{ALuck} programs be
``well-concretized": when a variable is concretized, all variables which occur
in constraints with it must also already have been concretized. This requirement
is strict, but it (a) makes it possible to run generators, and (b) simplifies
the generator inference process by shrinking the class of generators under
consideration.

@@

The fact that generators can fail is crucial: different generators for the same property can fail more or
less often, and we would like to infer generators which fail infrequently.
\begin{definition}[$p$-Goodness]
We say that a generator script $s$ is $p$-good if it succeeds with probability at least $p$.
\end{definition}

\begin{figure}
\caption{Generator Script Syntax}
\label{fig:aluck-syntax}
$$
\begin{array}{llll}
\text{Generator Actions} & a & ::= & !x \;|\; e R e\\
\text{Generator Scripts} & s & ::= & [] \;|\; a \, :: \, s\\
\end{array}
$$
\end{figure}

\begin{figure}
\caption{Generation Semantics}
\label{fig:aluck-semantics}
\begin{mathpar}
\inferrule{v \sim \mathcal{U}\left(\llbracket C(x)\rrbracket\right)\\ V\{x \mapsto v\},C \vdash s \Downarrow V',C'}{V,C \vdash !x :: s \Downarrow V',C'}

\inferrule{C'(x) = C(x) \cup \{c\} \\ V,C' \vdash s \Downarrow V',C'' }{V,C \vdash c(x) :: s \Downarrow V',C''}

\inferrule{V \vDash c \\ V,C \vdash s \Downarrow V',C'}{V,C \vdash c :: s \Downarrow V',C'}
\end{mathpar}
\end{figure}

@@

\section{Generator Learning with Bandits Algorithms}
\label{sec:bandits}
With our bag of generators in hand, we now need to find the best one. The
approach we take will be inspired by the Multi-Armed Bandits
\cite{gittins1979bandit} problem\footnote{More precisely, Stochastic Bernoulli
Bandits} from the theory of reinforcement learning. In short, the multi armed
bandits problem describes a situation where an algorithm is repeatedly presented
a fixed set of choices. Each choice gives a different (random) reward, and the
goal of the game is to maximize the total reward. Solving this optimization problem
online is too difficult in an adversarial setting, so algorithms for the multi-armed
bandits problem aim to instead minimize \emph{regret}: how much worse their total reward
was than the best possible total reward in hindsight. To achieve this, bandits algorithms
learn which actions have historically given more reward, and play those more frequently.
In our setting, the ``choices" are our
generator candidates, and the rewards are given by success or failure of a
generator to yield a value. Under this analogy, an algorithm for the Multi-Armed
Bandits problem will let us learn which generators give the best results
\textit{while simultaneously} generating a stream of valid inputs for the
function.

% To make the discussion more concrete, the Multi-Armed Bandits problem is
% described as the following repeated game: at each round $t$, the player plays an
% action $a_t \in [K]$, and receives a reward $X_{{a_t},t}$, which is a
% $\{0,1\}$-valued random variable. The random variables $X_{i,t}$ are IID for
% fixed $i$, and independent for fixed $t$. We write the mean of the $i$th reward
% variable (for all $t$) $\mu_i$. The goal of the game is to maximize one's
% reward, and so the goal of a bandits learning algorithm is to learn an adaptive
% policy, which takes a history of play up until state $t$ (all actions $a_{t'}$
% and received rewards $X_{a_{t'},t'}$ for $t' < t$), and produces a new action
% $a_t$. The metric by which we compare bandits algorthms is \textit{regret}: how
% much worse they do than the best policy in hindsight.

% \begin{definition}[Regret]
% Define $i_\star = \argmax_i \mu_i$, and write $\mu_\star = \mu_{i_\star}$. The regret $R(A)$ of an algorithm $A$ over $T$ rounds is defined as
% $$
% R(A) = T\mu_\star - \mathbb{E}\left[\sum_{t=1}^T X_{A(t),t}\right]
% $$
% \end{definition}

The algorithm we will use is called UCB1 \cite{auer2002finite}. While more sophisticated
bandits algorithms exist, this one is sufficient for our purposes. When given a list of generators
$g_1,\dots g_k$ and a number of rounds $T$ to run for, UCB1 emits some number $m \leq T$ of values generated by the
generators it was given. The main upshot is that UCB1 rapidly learns which generator among the ones it was given
fails the \emph{least often}, and so the ratio success ratio $\frac{m}{T}$ of this derived generator should, in the limit,
be no worse than the best generator.

% UCB1 achieves a
% regret bound of $O(\sqrt{KT\log K})$. While algorithms are known which achieve
% $O(\sqrt{KT})$ regret, UCB1 is simple to implement and understand, and performs
% more than well enough for our setting. In \autoref{alg:ucb1}, we present the
% UCB1 algorithm, modified for use in property-based testing. The algorithm takes
% generator candidates $g_1,\dots g_k$, the property $P$, a number of rounds to
% run for $T$, and returns a list of generated values satisfying $P$.

\begin{restatable}{thm}{regretthm}
Fix a property $P$ and generators $g_1,\dots,g_K$ which are $p_i$-good. Then,

$$
\lim_{T \to \infty} \mathbb{E}\left[\frac{1}{T}\left|\texttt{ucb1}(g_1,\dots,g_K,P,T)\right|\right] \geq \max_i p_i
$$
\end{restatable}

% Intuitively, this theorem justifies our intuition that the ucb1 algorithm
% ``finds the best generator". As the number of rounds gets large, the ratio of
% the number of valid inputs generated by \texttt{ucb1} to the number of rounds is
% no worse than the success probability of the best generator. One can also think
% of this theorem as saying that, with high probability\footnote{One could do an
% analysis using Chebyshev's inequality to bound the difference between the
% empirical frequency of the sequence and its mean.}, that the stream of values
% emitted by an ``infinite run" of \texttt{ucb1} acts like a $(\max_i p_i)$-good
% generator for $P$.

@@

I have implemented a prototype of a deductive verifier backed by PBT based on the above algorithms, it is available \href{https://github.com/jdublu10/triple-testing}{here}. All told, it is approximately 1000 lines of Haskell, approximately 300 lines of which are parsing code. The implementation makes use of the QuickCheck library \cite{qc} for generation combinators. The verifier implements a variant of the \texttt{IMP} language \cite{imp}
%https://dl.acm.org/doi/10.1145/361953.361966
with assignment, conditionals, and iteration. The only novel aspect of the implementation is the generation algorithm described in Sections 3,4, and 5: the testing of Hoare triples is trivial to implement, and is essentially folklore at this point.

Because the generator inference algorithm implemented by the prototype is the first of its kind, there are no existing tools we can compare it to in a benchmark. Instead, we simply present our results as a table (\autoref{tab:data}) of data about runs of the algorithm on different inputs. We leave it to the reader to decide if these results are impressive.

In \autoref{tab:data}, we show the results for some trials of our algorithm. In each trial, we run the generration algorithm on the specified constraint five times for $T = 1000$ iterations. $K$ is the average number of unique scripts discovered by the random DFS, Rejected is the average number of failed generator samples that happen during the runs, and Unique is the average number of unique input sets generated. In all cases, the generation takes less than 1ms of wall clock time, as reported by the Haskell library \texttt{timeit}. By default, integers are generated in the range $[-10000,10000]$ instead of the full integer range.

As \autoref{tab:data} indicates, more work is required in order for this algorithm to be ready for prime time. Most of this work needs to be done on the script inference side of things, to get better generator candidates. One low hanging fruit is a range analysis. This is exemplified by the (huge) performance differences between the runs on the proposition $0 \leq x \leq y \leq 100$ with and without the added constraint $x \leq 100$. This constraint is already implied by the first, but not explicitly including it degrades performance significantly. The ``best" generator in for the first proposition samples an arbitrary $x \geq 0$, which is very unlikely to leave room for some $y$ satisfying $x \leq y \leq 100$. Including the added constraint (which can be inferred from the first) improve performance dramatically. Another source of potential improvement is in \autoref{alg:generate-ordering}, where we could attempt to remove edges from the graph in order to minimize the number of scripts. Finally, we have only scratched the surface of reinforcement learning in \autoref{alg:ucb1}. Further tricks such as a probabilistic policy and dropping particularly bad generators could lead to faster convergence of the learner.

\begin{center}
\begin{table}[]
\caption{Results}
\label{tab:data}
\centering
\begin{tabular}{|l|l|l|l|l|}
\hline
Constraint & $K$ & Rejected & Unique \\ \hline
     $0 \leq x \leq y \leq 100 \wedge x \leq 100$ & $2$ & $11$ & $831$ \\ \hline
     $x = y + z \wedge x,y,z \geq 0$      &   $5$   &    $506$      &     $494$   \\ \hline
     $x = y + z \wedge x \geq y \wedge x,y,z \geq 0$      &   $4$  &    $48.6$      &    $951.4$    \\ \hline
     $0 \leq x \leq y $  &  $2$   &  $24.6$   &       $975.4$       \\ \hline
     $0 \leq x \leq y \leq 100$ & $2$ & $993.4$ & $6.6$ \\ \hline
     $0 \leq x \leq y \leq 100 \wedge x \leq 100$ & $2$ & $11$ & $831$ \\ \hline
     $x - y \leq 5 \wedge x - y \leq 10 \wedge y - z \leq 2$ & $3.8$ & $26.2$ & $831$ \\ \hline
\end{tabular}
\end{table}
\end{center}

\bibliographystyle{plain}
\bibliography{refs}

\newpage
\appendix
\section*{Appendix}

@@

>>>>

\usepackage{lipsum}  
\usepackage{graphicx}               % Necessary to use \scalebox

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Property-Based Testing in the style of QuickCheck has proven to be a
  very powerful and useful generalization of traditional software testing techniques
  such as unit testing. This power comes at the cost of requiring users to occasionally write \emph{generators}:
  programs which emit random data satisfying the invariants assumed by the program under test. While some work exists
  to derive these generators directly from the invariants in question, this work is mostly focused on highly structured data,
  and often fails to handle the kinds of numerical invariants that occur commonly in systems programming tasks.
  In this extended abstract, we present an approach to inferring generators for a restricted class of numerical properties.
  \vspace{-18px}
\end{abstract}

\maketitle

\section{Motivation}
% \vspace{-3px}
Property-Based Testing (PBT) \cite{qc} is a program testing method which tests
logical properties of functions by randomly generating thousands of inputs, and ensuring that
the the properties hold at each concrete value. Given a property $\forall x. P(x) \implies Q(x)$,
we generate values satisfying $P$, and check that they satisfy $Q$.
While this process is highly automated and requires no user input, users must occasionally write \emph{generators}:
programs which emit random data that satisfies the premise $P$.

Much of the time, this burden can be mitigated by automation, such as when the data structure being generated
is composed entirely of structures that already have generators, or $P$ is described as an inductive
relation \cite{gggir}. However, the case of inferring generators for \emph{numerical} properties---those involving constriants over numbers--- is not well
handled by prior work. Numerical properties appear regularly in
systems programming tasks \cite{???}, where structured data described by inductive invariants is rare, inputs regularly
take the form of pointers and integers, and many function preconditions amount to bounds-checking or ensuring
that certain fields in different structures are related in some way.

In this extended abstract, we develop a method for inferring random generators for a small class of numerical properties akin
to the kind commonly found in systems programming. Because manual generator-writing for numerical properties is especially
tedious, our goal is to infer generators that are comparable to ones that a user might write as a first cut.
The method proceeds by developing a number of generator ``candidates'', expressed in a DSL we call \texttt{ALuck}, described in \autoref{sec:luck}.
These candidates mimic the way that a human might write a generator for numerical preconditions: by randomly choosing values
one variable at a time, constraining subsequent choices by the previous values chosen. In \autoref{sec:sci}, we show how candidates
are chosen from the space of possible \texttt{ALuck} generators.
As we will see, it is sometimes the case that \emph{none} of the candidates are all that good. In this case, we must make the best
of a bad situation, and discover the best candidate among the set. A learning-theoretic method for this is presented in \autoref{sec:bandits}.
Finally, in \autoref{sec:impl}, we describe our implementation of
this algorithm, present benchmarks, and briefly discuss future work.

@@

The syntax of propositions that we handle are shown in
\autoref{fig:precond-syntax-and-aluck}. These are conjunctions and negations of
arithmetical equalities and inequalities with variables, which range over
the arguments of the function under test. The allowable expressions in these inequalities are
essentially multi-linear functions, where each variable occurs with degree at
most one. While this form is restrictive, we have found empirically that this covers
a wide range of preconditions in systems verification. Moreover, there
appears to be no inherent difficulty to extending the technique to handle
other numerical operators (div, mod), and constraints over structured types like lists: we present the algorithm as-is to
simplify the discussion.

\begin{figure}
\vspace{-10px}
\caption{Syntax of Propositions and Generator Scripts}
\label{fig:precond-syntax-and-aluck}
\scalebox{0.8}{\parbox{\linewidth}{%
$$
\begin{array}{llll}
\text{Expressions} & e & ::= & x \; | \; e + e \; | \; -e \;|\; e*e\\
\text{Relations} & R & ::= & = \;|\; \leq \;|\; <\\
\text{Propositions} & P & ::= & P \wedge P \;|\; e R e\\
\end{array}
$$
}}

\scalebox{0.8}{\parbox{\linewidth}{%
$$
\begin{array}{llll}
\text{Generator Actions} & a & ::= & !x \;|\; e R e\\
\text{Generator Scripts} & s & ::= & [] \;|\; a \, :: \, s\\
\end{array}
$$
}}
\vspace{-15px}
\end{figure}

To make the generator inference problem tractable, we fix the
syntactic form of the generators. Our generators are written
in a language called \texttt{ALuck} (for \textit{arithmetic} \texttt{Luck})
inspired by \texttt{Luck} \cite{luck}, a domain-specific language for writing
generators. Generators written \texttt{ALuck} run by sequentially \textit{constraining} and
\textit{concretizing} variables. Every variable in an \texttt{ALuck} generator
begins as a symbolic variable. Constraints over these variables are then added.
Variables can then be concretized, wherein they are replaced by a random value
satisfying the constraints on that variable that have been accumulated thusfar.
The final result of the generator is a map from variables to their (randomly)
chosen values.

More concretely, generators in \texttt{ALuck} are sequences of ``concretize"
operations, written $!x$, and ``constrain" operations, written simply as the
constraint to be added. This syntax is shown in
\autoref{fig:precond-syntax-and-aluck}. These sequences are then evaluated from left to
right while maintaining a pair of mappings: one from concretized variables to
their values, and the other from yet-unconcretized variables to the set of
constraints that have accumulated on them.  When a ``constrain" operation $c$ is
encountered, the constraint $c$ is added to the constraint sets of all of the
variables it mentions. If a constraint mentions no variables, it is checked for
valdity: if the constraint is not valid, the generator fails.

When a ``concretize" operation $!x$ is encountered, a value is randomly sampled from
the uniform distribution on the set of possible values\footnote{ Because
integers are bounded by machine word lengths, the ``uniform distribution" does
make sense here, even for unconstrained variables.  } denoted by the constraints
on $x$. This semantics is shown in \autoref{fig:aluck-semantics}, where the
judgment $V,C \vdash s \Downarrow V',C'$ means that the script $s$ evaluates
under concretized-variable-map $V$ and constraint map $C$, and returns updated
variable and constraint maps $V'$ and $C'$. We note that this semantics is partial,
as generators can fail when attempting to concretize a variable whose constraints are unsatisfiable.
This is crucial: different generators for the same property can fail more or
less often, and we would like to infer generators which fail infrequently.

To make the sampling step tractable, we enforce that our \texttt{ALuck} programs be
``well-concretized": when $!x$ occurs, all variables which occur
in constraints with $x$ must already have been concretized.

@@

\begin{figure}
\vspace{-10px}
\caption{Generation Semantics}
\label{fig:aluck-semantics}

\scalebox{0.75}{\parbox{\linewidth}{%
\begin{mathpar}
\inferrule{v \sim \mathcal{U}\left(\llbracket C(x)\rrbracket\right)\\ V\{x \mapsto v\},C \vdash s \Downarrow V',C'}{V,C \vdash !x :: s \Downarrow V',C'}

\inferrule{C'(x) = C(x) \cup \{c\} \\ V,C' \vdash s \Downarrow V',C'' }{V,C \vdash c(x) :: s \Downarrow V',C''}
\hspace{5px}
\inferrule{V \vDash c \\ V,C \vdash s \Downarrow V',C'}{V,C \vdash c :: s \Downarrow V',C'}
\end{mathpar}
}}
\vspace{-15px}
\end{figure}

@@

\section{Generator Learning with Bandits Algorithms}
\label{sec:bandits}
With our bag of generators in hand, we now need to find the best one. The
approach we take will be inspired by the Multi-Armed Bandits
\cite{gittins1979bandit} problem from the theory of reinforcement learning. In short, the multi armed
bandits problem describes a situation where an algorithm is repeatedly presented
a fixed set of choices. Each choice gives a different (random) reward, and the
goal of the game is to maximize the total reward. Solving this optimization problem
online is too difficult in an adversarial setting, so algorithms for the multi-armed
bandits problem aim to instead minimize \emph{regret}: how much worse their total reward
was than the best possible total reward in hindsight. To achieve this, bandits algorithms
learn which actions have historically given more reward, and play those more frequently.
In our setting, the ``choices" are our
generator candidates, and the rewards are given by success or failure of a
generator to yield a value. Under this analogy, an algorithm for the Multi-Armed
Bandits problem will let us learn which generators give the best results
\textit{while simultaneously} generating a stream of valid inputs for the
function.

The algorithm we will use is called UCB1 \cite{auer2002finite}. While more sophisticated
bandits algorithms exist, this one is sufficient for our purposes. When given a list of generators
$g_1,\dots g_k$ and a number of rounds $T$ to run for, UCB1 runs the generators $g_i$ in a way that attempts
to learn which generator succeds the most frequently, while also attempting to not waste time by running
failing generators too much. In the end, UCB1 emits the $m \leq T$ successfully-generated values over the course of its run.
The main upshot is that the ratio success ratio $\frac{m}{T}$ of this derived generator should, in the limit,
be no worse than the \emph{best} generator the algorithm was given. 
% UCB1 achieves a
% regret bound of $O(\sqrt{KT\log K})$. While algorithms are known which achieve
% $O(\sqrt{KT})$ regret, UCB1 is simple to implement and understand, and performs
% more than well enough for our setting. In \autoref{alg:ucb1}, we present the
% UCB1 algorithm, modified for use in property-based testing. The algorithm takes
% generator candidates $g_1,\dots g_k$, the property $P$, a number of rounds to
% run for $T$, and returns a list of generated values satisfying $P$.
\begin{restatable}{thm}{regretthm}
Fix a property $P$ and generators $g_1,\dots,g_K$, where $g_i$ succeeds with probability at least $p_i$. Then,
\vspace{-2px}
\[
\lim_{T \to \infty} \mathbb{E}\left[\frac{1}{T}\left|\texttt{ucb1}(g_1,\dots,g_K,P,T)\right|\right] \geq \max_i p_i
\]
\end{restatable}
% Intuitively, this theorem justifies our intuition that the ucb1 algorithm
% ``finds the best generator". As the number of rounds gets large, the ratio of
% the number of valid inputs generated by \texttt{ucb1} to the number of rounds is
% no worse than the success probability of the best generator. 
In other words, with high probability
% \footnote{One could do an
% analysis using Chebyshev's inequality to bound the difference between the
% empirical frequency of the sequence and its mean.}
, the stream of values
emitted by an ``infinite run" of UCB1 acts like a $(\max_i p_i)$-good
generator for $P$.

@@

I have implemented the generator inference algorithm described in this abstract,
as well as a deductive program verifier (a la Dafny or Frama-C \cite{??}) for the
\texttt{IMP} language which uses these generators as a verification backend.
The code is available \href{https://github.com/jdublu10/triple-testing}{here}.

% Because the generator inference algorithm implemented by the prototype is the
% first of its kind, there are no existing tools we can compare it to in a
% benchmark. Instead, we simply present our results as a table
% (\autoref{tab:data}) of data about runs of the algorithm on different inputs. We
% leave it to the reader to decide if these results are impressive.

In \autoref{tab:data}, we show the results for some trials of our algorithm. In
each trial, we run the generration algorithm on the specified constraint five
times for $T = 1000$ iterations. 
% $K$ is the average number of unique scripts
% discovered by the random DFS, Rejected is the average number of failed generator
% samples that happen during the runs, and Unique is the average number of unique
% input sets generated. In all cases, the generation takes less than 1ms of wall
% clock time. Integers are generated in the range $[-10000,10000]$ instead of the full integer
% range.

In \autoref{tab:data-selected}, we present selected results: a larger table
is available in the Appendix. The largest issue is the performance
differences between the runs on the proposition $0 \leq x \leq y \leq 100$ with
and without explicitly including the implicand $x \leq 100$.
The ``best" generator for the proposition without it samples an
arbitrary $x \geq 0$, which is very unlikely to leave room for some $y$
satisfying $x \leq y \leq 100$. Including the added constraint ensures that we leave space
with our initial choice of $x$.

% Further, we have only scratched the surface of reinforcement learning
% in \autoref{alg:ucb1}. Further tricks such as a probabilistic policy and
% dropping particularly bad generators could lead to faster convergence of the
% learner.

\begin{table}[h]
\label{tab:data-selected}
\centering
\scalebox{0.8}{\parbox{\linewidth}{%
\begin{tabular}{|l|l|l|l|l|}
\hline
Constraint & $K$ & $R$ & $U$ \\ \hline
    %  $0 \leq x \leq y \leq 100 \wedge x \leq 100$ & $2$ & $11$ & $831$ \\ \hline
     $x = y + z \wedge x \geq y \wedge x,y,z \geq 0$      &   $4$  &    $48.6$      &    $951.4$    \\ \hline
     $0 \leq x \leq y \leq 100$ & $2$ & $993.4$ & $6.6$ \\ \hline
     $0 \leq x \leq y \leq 100 \wedge x \leq 100$ & $2$ & $11$ & $831$ \\ \hline
     $x - y \leq 5 \wedge x - y \leq 10 \wedge y - z \leq 2$ & $3.8$ & $26.2$ & $831$ \\ \hline
\end{tabular}
\caption{Selected Results. $K$ = average number of unique scripts, $R$ = number of failed draws, $U$ = number of unique values generated}
}}
\end{table}

\clearpage
\bibliographystyle{plain}
\bibliography{refs}

\clearpage
\appendix
\section*{Appendix}

\begin{center}
\begin{table}[h]
\label{tab:data-full}
\centering
\scalebox{0.8}{\parbox{\linewidth}{%
\begin{tabular}{|l|l|l|l|l|}
\hline
Constraint & $K$ & $R$ & $U$ \\ \hline
     $0 \leq x \leq y \leq 100 \wedge x \leq 100$ & $2$ & $11$ & $831$ \\ \hline
     $x = y + z \wedge x,y,z \geq 0$      &   $5$   &    $506$      &     $494$   \\ \hline
     $x = y + z \wedge x \geq y \wedge x,y,z \geq 0$      &   $4$  &    $48.6$      &    $951.4$    \\ \hline
     $0 \leq x \leq y $  &  $2$   &  $24.6$   &       $975.4$       \\ \hline
     $0 \leq x \leq y \leq 100$ & $2$ & $993.4$ & $6.6$ \\ \hline
     $0 \leq x \leq y \leq 100 \wedge x \leq 100$ & $2$ & $11$ & $831$ \\ \hline
     $x - y \leq 5 \wedge x - y \leq 10 \wedge y - z \leq 2$ & $3.8$ & $26.2$ & $831$ \\ \hline
\end{tabular}
\caption{Full Results}
}}
\end{table}
\end{center}
\noindent\rule{0.5\textwidth}{1pt}

@@

\subsection*{Bandits and UCB1}
The Multi-Armed Bandits problem is
described as the following repeated game: at each round $t$, the player plays an
action $a_t \in [K]$, and receives a reward $X_{{a_t},t}$, which is a
$\{0,1\}$-valued random variable. The random variables $X_{i,t}$ are IID for
fixed $i$, and independent for fixed $t$. We write the mean of the $i$th reward
variable (for all $t$) $\mu_i$. The goal of the game is to maximize one's
reward, and so the goal of a bandits learning algorithm is to learn an adaptive
policy, which takes a history of play up until state $t$ (all actions $a_{t'}$
and received rewards $X_{a_{t'},t'}$ for $t' < t$), and produces a new action
$a_t$. The metric by which we compare bandits algorthms is \textit{regret}: how
much worse they do than the best policy in hindsight.

\begin{definition}[Regret]
Define $i_\star = \argmax_i \mu_i$, and write $\mu_\star = \mu_{i_\star}$. The regret $R(A)$ of an algorithm $A$ over $T$ rounds is defined as
$$
R(A) = T\mu_\star - \mathbb{E}\left[\sum_{t=1}^T X_{A(t),t}\right]
$$
\end{definition}
